Algebraic Groups\vspace{-1ex}

### 2020-01-17 Algebraic Geometry/Algebraic-Geometry.tex
Update

\section{September 25, 2019}
The first thing that one asks is ``what is geometry?'' One needs to be able to answer this 
question before they define AG. One idea is that geometry is topology + structure.

@@

>>>>

\part{Quarter 1: Spaces with functions}
\section{September 25, 2019}
The first thing that one asks is ``what is geometry?'' One needs to be able to answer this 
question before they define AG. One idea is that geometry is topology + structure.

@@

\part{Quarter 2: Sheaves and Schemes}
I missed the first five lectures of this quarter because I was travelling for a conference. Sorry about that!

\section{January 17th, 2020}
\subsection{SCHEMES}
Let's just dive right in. Last time they were talking about things:
\begin{itemize}
	\item Is the category of spaces with functions a full subcategory of locally ringed spaces?
	\item Thinking about $\Hom(\Spec A,\O_A),(\Spec B,\O_B))$
\end{itemize}
In the later case, we noticed that if we had a pair of maps $(f,f^\sharp)$ between $\Spec$s and the structure sheafs. Then $\Gamma(f^\sharp):B\to A$
gives us a local condition: if $\varphi=\Gamma(f^\sharp)$, $f$ must be the map induced by $\varphi$.

\subsection{Talking about affine schemes}
Let $A$ be a (commutative unital) ring. Define a locally ringed space $\Spec A$ where 
\begin{itemize}
	\item The underlying space $|\Spec A|$ is the set of prime ideals of the ring.
	\begin{itemize}	
		\item Notice before we used the maximals but getting all the primes gives us something extra...
		\item The topology is the Zariski topology.
		\item This looks like a closed point for every prime in $\bbZ$ as well as the dense generic point $(0)$. This is the French model for the Italian idea of a ``general point.''
		\item A basis for this topology is $D(f)\subseteq |\Spec A|$ defined as the nonvanishing locus:
		\[D(f)=\{\frakp|f\notin\frakp\}\]
		or in other words, $f(\frakp)\ne 0$.
	\end{itemize}
	\item We want 
	\[\Gamma(|\Spec A|,\O)=\O(|\Spec A|)=A.\]
	as well as $\O(D(f))=A_f$, the localization. We know that $D(f)$ defines a basis for the topology for $\Spec A$, so this implies
	this gives us a way to define the sheaf locally.

	So it suffices to show that $\O:\calB\op\to\Ring$ forms a sheaf (satisfies the sheaf condition). Here's the (``bestest'') proof:
	\begin{prf}[Bestest]
		Algebraic formulation: Begin by reducing to $U=|\Spec A|$. Then take an open cover $U_i$ of $U$. Look at descent theory! It's also in SGA.
	\end{prf}
\end{itemize}
\begin{rmk}
	We can think of the value of $f$ at a prime $\frakp$ as follows: $f\in A$ is an element of a ring and 
	\[f(\frakp)\eqdef \bar f\in \kappa(A/\frakp)\]
	the residue field.
\end{rmk}
\begin{rmk}
	To see why it suffices to define the sheaf on a basis: let $\calB$ be a basis for a topological space $X$. Then let $F:\calB\op\to \Ab$ be a sheaf 
	on $\calB$. Then there is some work to do, but the statement is that the category of sheaves on $X$ is equivalent to the category of sheaves on $\calB$.

	The idea is as follows: to get essential surjectivity, you take a sheaf on $\calB$ and take the sheafification to get a sheaf on $X$. 
\end{rmk}

### 2020-01-17 Algebraic Groups/Algebraic-Groups.tex
Update

\section{September 25, 2019}
\subsection{Group objects}
Let $\calC$ be a category with a final object and finite products.
\begin{defn}
	A \textbf{group object $G$ in $\calC$} is an object in $\calC$ along with multiplication, identity, and inverse morphisms
	satisfying the usual axioms.
\end{defn}

@@

>>>>

\part{Quarter 1: Structure Theory}
\section{September 25, 2019}
\subsection{Group objects}
Let $\calC$ be a category with a final object and finite products.
\begin{defn}
	A \textbf{group object $G$ in $\calC$} is an object in $\calC$ along with multiplication, identity, and inverse morphisms
	satisfying the usual axioms.
\end{defn}

@@

\part{Quarter 2: Representation Theory}
\section{January 17th, 2020}
\subsection{Recap}
Let $T\subseteq B\subseteq G$ be a maximal torus, Borel, and semisimple algebraic group. Then we talked about:
\begin{itemize}
	\item The infinitesimal theory by studying $\Lie G$
	\item Borel subgroups, parabolic subgroups, and the flag variety $G/B$.
	\item We also saw the categorization of split connected semisimple algebraic groups according to their root data:
	This is either given by $(X,R,X^\vee,R^\vee)$ of a group $X$ and a root system $R$ and their duals or equivalently a 
	weight lattice $\Lambda$ and root latice $\Lambda_r=\bbZ R$. Then we can define 
	\[Z/\bbZ R\cong \Lambda/\Lambda_r\]
	to be the \textbf{fundamental group.}
	\item Representation theory: $\Rep G$ is what is called a \textbf{highest weight category.} Let $\Lambda=X$. Then 
	we have the \textbf{dominant roots}
	\[\Lambda_+=\{\lambda\in\Lambda|\langle\lambda,\alpha^\vee\rangle\ge 0,\alpha\in R^+\}\]
	and for every dominant root $\lambda$, we have the \textbf{(co)standard} modules $\Delta(\lambda)$ and $\nabla(\lambda)$.
\end{itemize}
\begin{defn}
	Let $\lambda\in\Lambda=\Hom(T,\Gm)$. Define the $T$-module $\calR_\lambda$ by 
	\[t\cdot 1\eqdef \lambda(t)1.\]
	Then $T(k)=(k^\times)^\Gamma$ gives us $\lambda(t)\in k^\times$.
\end{defn}

\begin{ex}
	Let $G=\GL_n$ and $B$ the upper triangular matrices and $T$ the diagonal ones. Then the unipotent matrices $\calU$ of upper triangulars with ones on the diagonal 
	include into $B$ and we have a section $T\to B$ giving us $B=T\rtimes U$.

	Then we consider $k_\lambda$ as a representation of $B$ by having $\calU$ act trivially. Then we define 
	\[\nabla(\lambda)=\Ind_B^G k_\lambda\]
	where $\Ind_B^G$ is the right adjoint to the restriction functor $\Res^G_B$.
\end{ex}
\begin{rmk}
	The induction functor always exists, but the \textit{left} adjoint to $\Res$ doesn't. When it does, it is called \textbf{coinduction.}
\end{rmk}

\subsection{The Associated Sheaf}
Given $H\subseteq G$, there is a functor
\[\calF:\Rep H\to \Sch G/H.\]
this leads to a fact:
\begin{prop}
	\[\Ind_B^G k_\lambda=\nabla(\lambda)\cong H^0(G/B,\calF_{G/B}(k_\lambda))\]
\end{prop}

The standard module $\Delta(\lambda)$ is essentially the dual (after applying an element of the Weyl group of longest length).
\begin{thm}
	Let $\lambda\in\Lambda_+$. Then $\nabla(\lambda)$ has a simple socle, denoted $L(\lambda)$. Furthermore, $\{L(\lambda)|\lambda\in\Lambda_+\}$ is a complete set of $G$-modules.
\end{thm}

\subsection{Other things to focus on:}
Look at Humphreys' \textit{Linear Algebraic Groups}. There is an appendix on root systems that is

\subsection{Examples}
Let $k=\bar k$ and $G=G(\bar k)$. For type $A_{n+1}$, we have $\SL_n$ as well as $\operatorname{PGL}_n$. In the former case, $\Lambda/\Lambda_r=\bbZ/n\bbZ$ and in the latter this quotient is trivial.

For type $C_n$, we get $\operatorname{Sp}_{2n}$. Let $\Sigma$ be the matrix given by $-1$s and 1s on the antidiagonal and zeros elsewhere (negative in the ``third quadrant.'')
Then the matrices in this are those preserving the associated form: that is 
\[A^T\Sigma A=\Sigma\]

For type $B_n$, we look at the $2n+1$ square matrix $\Sigma$ that is block diagonal with blocks $(1)$ and the previous matrix. Then $\operatorname{SO}(2n+1,k)$ are the matrices preserving $\Sigma$.
The type $D_n$ matrices are the ones preserving the matrix given by (positive) ones on the antidiagonal.

\subsection{More Generally...}
Let $A$ be a separable associative unital algebra over $k$. (That is, $A_{\bar k}$ is a product of simple algebras). Fix an involution $\sigma:A\to A$ 
($\sigma^2=\id$). There is always an associated bilinear form for these involutions, which can be symmetric or skew-symmetric. It either preserves the base field or else 
it descends to an involution of this field.

\begin{defn}
	$\GL_{1,A}$ is a group scheme whose $R$-points ($R$ a ring) are given by 
	\[\GL_{1,A}(R)=\left((A\otimes_k R)^\times\right)^k\]
	which contains a subgroup scheme 
	\[\operatorname{Iso}(A,\sigma)(R)\eqdef \{a\in(A\otimes_k R)^\times|a\cdot\sigma_R(a)=1\}\]

	We also have $\Aut(A,\sigma)\subset \Aut(A)$ defined as 
	\[\Aut(A,\sigma)(R)\eqdef\{\alpha\in\Aut_R(A_R)|\alpha\circ\sigma_R=\sigma_R\circ\alpha\}\]

	And finally,
	\[\operatorname{Sim}(A,\sigma)\eqdef \{a\in A_R^\times|a\cdot\sigma_R(a)\in k_R^\times\}\]
\end{defn}

Now if $A$ is a central simple algebra and $\sigma$ is a symplectic form, then 
\begin{itemize}
	\item $\operatorname{Sp}(A,\sigma)=\operatorname{Iso}(A,\sigma)$
	\item $\operatorname{GSp}(A,\sigma)=\operatorname{Sim}(A,\sigma)$
	\item $\operatorname{PGSp}(A,\sigma)=\Aut(A,\sigma)$
\end{itemize}

Want to know more? Look at \textit{The Book of Involutions.}
